---
title: "Gov 50 Final Project"
author: "Adamari Cruz"
description: "My final project on crane efficiency in US ports (2016-2020)"
output:
  distill::distill_article:
    self_contained: false
---



## Introduction

In this project, I explore data related to port infrastructure. More specifically, this investigation looks at port efficiency in terms of shipping containers. The shipping container was one of the most important inventions of the 20th century, but now we are confronted with a shipping container shortage because they are just sitting in ports, waiting to be unloaded/loaded. COVID left us with a supply chain crisis, and this won't be solved until the containers are more quickly unloaded and turned around. Because of this pressing issue, my project examines cranes and crane efficiency, the machines that load and unload all the shipping containers at ports. 

My research question: Is there a relationship between the efficiency of a port and the number of shipping container cranes? In this study, I plan to examine the extent to which the number of cranes in a given port  affects the efficiency (how many shipping containers are imported and exported in a year). While ports are normally categorized by cargo volume, I'm curious about crane capacity. I hypothesize that ports with more cranes are more efficient, since they can unload and load vessels at a higher rate. If I observe that an increase in efficiency as the number of cranes increases, then this would support my hypothesis. On the other hand, if there's a decrease in efficiency as number of cranes stays the same or increases, then this would suggest that my hypothesis is false. This study would be insightful for sustainability and climate conscious reasons, as expanding ports is costly and can be harmful to the environment. We would have to find creative solutions to support the ever increasing networks of globalization, such as cranes that take up less space and are more versatile. 
  
## Data and Research Method

I analyze data about port volume in terms of imports and exports in the US from the years 2016-2020, not including 2017, because one year may be an anomaly. Because I consider 2016-2020 to represent one point in time, this is a cross-sectional design. My data set pulls information from two separate data sets, one that focuses on cranes and the other on port efficiency. Because the data sets had overlapping information about some of the same ports, I combined the data into a master "joined" set to only include information on the ports that were present in both data sets. 


In this master data set, I include port ID, year, trade type, percent change, units, number of cranes, port (name), and crane change. The most important variables here are percent change, number of cranes, and crane change. Percent change is my measure of efficiency, as it measures the change of volume of shipping containers TEU's imported and exported at a port from the previous year. Number of cranes describes itself, and crane change is the percent change in number of cranes during the 4 years period. This is a variable I created by calculating the difference between the max and min values for the number of cranes at a given port, dividing by the min value for the number of cranes, and multiplying by 100. Since the dependent variable, percent change, is looked at from the crane perspective, the following plot paints a better picture of what the data reveals.

```{r}
library(tidyverse)
library(dbplyr)
library(ggplot2)
library(broom)

crane_data<-read_csv("data/realcrane_data.csv")

port_data<-read_csv("data/port_data_UPDATED.csv")

port_data<-rename(port_data, Year = 'Reporting Year')
port_data$Year <- as.numeric(as.character(port_data$Year))

port_data<-port_data|>
  filter(Year==c(2016,2018,2019))

joined<- port_data|>
  left_join(crane_data)|>
  select(Year, `Port ID`, `Trade Type`, `Percent Change`, `Units`, `Number of Cranes`, `Port`)

joined<-joined|>
  group_by(`Port ID`)|>
  mutate(crane_change=((max(`Number of Cranes`)- min(`Number of Cranes`))/min(`Number of Cranes`))*100)|>
  drop_na()|>
  filter(`Trade Type`=="TOTAL")

joined$`Port ID`<- with(joined, reorder(`Port ID`, `Number of Cranes`))

plot1<-ggplot(data = joined, mapping = aes(x=`Number of Cranes`, y=Year))+
              geom_point(aes(colour = factor(`Port ID`)))+
  labs(title="Number of Cranes for Various Ports from 2016-2019", colour = "Port ID")
plot1

```

In this first plot, we can see an upward change in the number of cranes from 2016-2019. From this, we know that ports invest money into acquiring more cranes, hoping to be able to handle more shipping containers more quickly. In the next section, we will see how this pans out.

## Results

Before performing a regression, the following scatter plot visualizes the bivariate relationship between the percentage change of efficiency (dependent variable) and the percentage change of number of cranes (main independent variable). 


```{r}
plot2<-ggplot(data = joined, mapping = aes(x=crane_change, y=`Percent Change`))+
              geom_point(aes(colour = factor(Year)))+ 
  labs(x="Percentage Change of # of Cranes", y="Percentage Change of Efficiency", title="Growth of Cranes vs Efficiency from 2016-2019", colour = "Year")+
geom_smooth(method= "loess", se=FALSE)
plot2
#plot1 data visualization i can compare the growth (rate of change) of cranes to the change in efficiency as a scatterplot

```

This visualization shows that from our data set of ports in the US from 2016-2019, the optimal percentage change of number of cranes was approximately 75% for port efficiency. After that, there are decreasing marginal returns on port efficiency past 75%. Even at a 300% increase, meaning tripling the number of cranes at a given port, the increase in port efficiency is not that much bigger than the percentage change of efficiency at a 75% increase in cranes. These results may suggest that investments shouldn't necessarily go to buying new cranes, but rather be invested in other aspects of a port that contribute to efficiency, such as the number of port workers or the expansion of port size.

Moving on to the multivariate regression model, the following tables summarizes these findings. The model includes the percentage change of number of cranes and year.

```{r}
joined[is.na(joined) | joined=="Inf"] = NA
model<- lm(data = joined, `Percent Change` ~ crane_change + `Number of Cranes`)|>
  broom::tidy()|>
  select(term, est=estimate, se=std.error, pval=p.value)

knitr::kable(model, digits=3, col.names = c("Variable", "Coefficient Estimate", "Standard Error", "P-Value"))
```

This table shows that the coefficient of interest for crane change is 0.025, meaning that a one unit increase in crane change produces a 0.025 increase in port efficiency. However, seeing that the p-value is 0.408, this is not statistically significant. I don't believe the coefficient estimate to represent a causal effect, as there are many other variables that could come into play, such as the size of the port, the number of port workers on site, the kind of coast the port is on, or the contracts with shipping vessels. 


## Conclusion 

In conclusion, while the regression shows a positive correlation between an increase in the number of cranes and port efficiency, these results are not statistically significant. Furthermore, the scatter plot suggests that there may be diminishing returns after a certain point of acquiring more cranes. There are many limitations in this study, including lack of data about other possible confounders as I mentioned above, missing data within the existing data sets, and not having information on ports worldwide to extend the data set. A way to improve this study would be to gather this information to paint a better picture of what has happened with crane capacity and port efficiency in the 21st century as globalization has rapidly expanded.







